Board Game Recommendation

### 2015-03-02 paper.Rnw
Additions and revisions to paper
Revised the introduction, began the technical implementation. Next up is
work on Bibliography and revising the modeling section, as well as
adding a conclusion.

Board gaming, while being very popular among children, is a niche hobby beyond adolescence. Everyone may have grown up with Monopoly, but most people may have never heard of Citadels or Twilight Imperium. And as such, board gaming is a very difficult hobby to break into. ``Gateway Games" such as Settlers of Catan or Ticket to Ride offer a gradual transition between the ultra-popular and mechanics-lite games like monopoly, to more rules-heavy games like Puerto Rico. The most difficult part of the transition, for most people, is actually finding new games to play. New games are often expensive for people who may be apprehensive about trying new games (a copy of Settlers of Catan goes for \$37.80 on Amazon.com). For this reason, a method to recommend games to people would be invaluable.

A few methods already exist, however, their usefulness varies. Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it will recommend more games that a person may like. A pretty thorough analysis of it and its issues can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}. The bigget issue is that it cannot be tailored to an individual. You, an individual user, may have different preferences than the users as a whole, and it might be more useful to tailor recommendations to the individual.

A redditor identified this issue, and created a board game recoomender on the subreddit boardgamerecommender. You can ``summon" the recommender by posting your boardgamegeek username anywhere on reddit, preceded by ``/u/boardgamerecommender". This will prompt the bot to go search through your ratings of games on boardgamegeek, and then it will make a post with 4 lists of games: New games you may enjoy, older games you may enjoy, games you may find underrated, and games you may find overrated. This is a huge step forward for individuals that are looking for games, as they can now rate games they've played, and can then get recommendations for games they might like. However, there are a few drawbacks. The process has a few possible user inputs (such as playtime, number of players, category, mechanic, etc) that will narrow down the games that will be returned to the user. However, with each specification, a new recommendation will be generated, a process that is said to take 10-30 minutes (FIND A BETTER ESTIMATE FOR THIS). Also, there is no published methodology, so we cannot view how these recommendations are made.

This being said, these are only two board game recommendation systems out of the many out there. In this paper, we seek to build a system the will recommend board games to users based on their own recommendations as well as the recommendations of others.

@@

data scraped from boardgamegeek.com
scraped in 2 sections - games and ratings
pulled down whole pages due to extensive nature
after pages were pulled down, code was written to extract the pertinent information from each
test data set was dumped into database
database has a few tables - game key, player key, ratings, classifiers, other
the structure of this database allows for easy functionality

@@

\subsection{Model Fit}
The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.

@@

\subsection{Classifier Aggrigation}
Independent of the global effects, we need to give each user a score in each classification category. We will use 4 types of categorical variables, which will yield (FIND OUT HOW MANY CLASSIFIER VARIABLES). To do this, we can either use the ratings themselves, or the residuals from the global effects process. Let $c_{u,j}$ be the categorical rating for user $u$ in category $j$. Additionally, for each game $i$, there exists a set of classifiers $C_i$ that describe the aspects of the game in things such as mechanics, game family, categories, and subdomains.  Then for each user, we take all of the categorical values for each user and normalize them, so they should all have mean 0 and standard deviation of 1 for each user. Then, within each category, the average rating is calculated. Now, if a user hasn't rated anything in this category we can assign them a score of 0 in that category (what should amount to an average rating). These serve dual purposes: Firstly, we can assess what sorts of games a user rates higher than average most of the time, and game they rate lower than average most of the time. This will allow the algorithm to not only recommend specific games that have a high predicted rating, but will also allow for the user to find categories or mechanics that they enjoy as well. The other purpose is that we now have a larger dataset to compare people than we might have had before, as we are guarantedd to have (FIND OUT HOW MANY CLASSIFIER VARIABLES) for every user.

\subsection{k-nearest-neighbors}
After we aggregate on the classifiers, we can then find correlations between all of the users. We will then take the $k$ nearest neighbors, where the distance is given by $\rho_{i,k}$, the correlation between the $i$ and $k$th users. In this algorithm, $k$ will be chosen by a crossvalidation procedure that will be done on a user-by-user basis, as determining a global value for $k$ would be far too computationally expensive. After the optimal $k$ is found, then for each game the $k$ nearest neighbors' (who have rated the game) rating's are averaged using a weighted mean, with the weights being equal to $\frac{2}{1 + \rho_{i,k}}$.

\subsection{Back to Rating}
After the k-nearest-neighbors step, we need to "rewind" from the global effects fitting process. For each game, we'll need to add on the specific user effect, the effect for each game, and the overall game mean. This will then yield a rating for each game for each player!

\section{Implementation}

\subsection{Individualized Summaries}

\subsection{Group Summaries}

@@

%\end{multicols}
\bibliographystyle{asa}
\bibliography{references}

>>>>

Board gaming, while being very popular among children, is a niche hobby beyond adolescence. Everyone may have grown up with Monopoly, but most people may have never heard of Citadels or Twilight Imperium. And as such, board gaming is a very difficult hobby to break into. ``Gateway Games" such as Settlers of Catan or Ticket to Ride offer a gradual transition between the ultra-popular and mechanics-lite games like monopoly, to more rules-heavy games like Puerto Rico. For most people, the difficult part of the transition is actually finding new games to play. New games are often expensive for people who may be apprehensive about trying new games (a copy of Settlers of Catan goes for about \$40 on Amazon). For this reason, a method to recommend games to people would be invaluable.

A few methods of varying usefulness already exist. Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it will recommend more games that a person may like. A pretty thorough analysis of it (and its issues) can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}. The bigget issue is that it cannot be tailored to an individual. You, an individual user, may have different preferences than the userbase average, and it might be more useful to tailor recommendations to the individual.

Recommendations for the individual aren't easy to come by. There are a few automated systems out there (such as the boargamerecommender bot on reddit), but for the most part they are time intensive and often lack interactivity or any sort of information that would be useful to the user beyond a few single games that it might recommend.

In this paper, we seek to build a system the will recommend board games to users based on their own recommendations as well as the recommendations of others. This system will offer the user not only recommendations for a very large number of games, but it will also offer insights to the user about what sorts of games they like, not only scores for particular games.

@@

For data, we turn to \url{boardgamegeek.com}, a website that holds information, ratings, and reviews for over 75,000 games. Users can create an account and then rate games they've played as well as keep an inventory of games they own. All of this information is accessable and scrapable, and will be the data that drives our analysis.

There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section). The characteristics we're interested in fall into these categories:

\begin{center}
\begin{tabular}{|l|c|}
\hline
Category &  This covers the broad categories a game\\
& might fall into (ex.: Humor, Puzzle, Sports, etc.) \\
\hline
Family &  Some games a naturally part of a "family" \\
&  that share a name or common element \\
& (ex.: Animals:Bats, Ancient Wars Series, Hello Kitty)\\
\hline
Mechanic &  This is what mechanisms the game uses\\
& (ex.: Dice Rolling, Card Drafting, Worker Placement)\\
\hline
Subdomain & More general than Category, but loosely bins games\\
& (ex.: Family Games, Strategy Games, etc.)\\
\hline
\end{tabular}
\end{center}

The ratings for each game are on a scale from 1 to 10, with a score of 1 indicating a really bad game and a score of 10 being the best possible. Each of the ratings are paired with a unique user identification number as well as a game number. Combined with the classifiers data, we can provide very specific ratings for each individual user.

@@

\subsection{Model Structure}
The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.

@@

\subsubsection{Classifier Aggrigation}
Independent of the global effects, we need to give each user a score in each classification category. We will use 4 types of categorical variables, which will yield (FIND OUT HOW MANY CLASSIFIER VARIABLES). To do this, we can either use the ratings themselves, or the residuals from the global effects process. Let $c_{u,j}$ be the categorical rating for user $u$ in category $j$. Additionally, for each game $i$, there exists a set of classifiers $C_i$ that describe the aspects of the game in things such as mechanics, game family, categories, and subdomains.  Then for each user, we take all of the categorical values for each user and normalize them, so they should all have mean 0 and standard deviation of 1 for each user. Then, within each category, the average rating is calculated. Now, if a user hasn't rated anything in this category we can assign them a score of 0 in that category (what should amount to an average rating). These serve dual purposes: Firstly, we can assess what sorts of games a user rates higher than average most of the time, and game they rate lower than average most of the time. This will allow the algorithm to not only recommend specific games that have a high predicted rating, but will also allow for the user to find categories or mechanics that they enjoy as well. The other purpose is that we now have a larger dataset to compare people than we might have had before, as we are guarantedd to have (FIND OUT HOW MANY CLASSIFIER VARIABLES) for every user.

\subsubsection{k-nearest-neighbors}
After we aggregate on the classifiers, we can then find correlations between all of the users. We will then take the $k$ nearest neighbors, where the distance is given by $\rho_{i,k}$, the correlation between the $i$ and $k$th users. In this algorithm, $k$ will be chosen by a crossvalidation procedure that will be done on a user-by-user basis, as determining a global value for $k$ would be far too computationally expensive. After the optimal $k$ is found, then for each game the $k$ nearest neighbors' (who have rated the game) rating's are averaged using a weighted mean, with the weights being equal to $\frac{2}{1 + \rho_{i,k}}$.

\subsubsection{Back to Rating}
After the k-nearest-neighbors step, we need to "rewind" from the global effects fitting process. For each game, we'll need to add on the specific user effect, the effect for each game, and the overall game mean. This will then yield a rating for each game for each player!

\subsection{Parameter Optimization \& Selection}
There are three parameters that the model relies on: The two $\alpha$ values that balance out biasedness in the global effects section, and the $k$ in the k-nearest neighbors portion of the algorithm. The $\alpha$'s will be done globally (that is, every individual will use the same values for $\alpha_1$ and $\alpha_2$), while the values for $k$ will be done locally (each user will have their own value of $k$). While the approaches for finding the optimal values for these parameters may be slightly different, the general implementation will be the same. We'll first begin by partitioning a subset of the data into 10 different sections (10 'folds'), and then choosing an appropriate parameter space for the parameter to be optimized. Ffor the alpha's, we'll choose values between 0 and 1000 (double check this), evenly spaced, and for $k$ we'll choose a positive integer. We'll then proceed fold by fold, using the other 9 folds to fit a model, and then using the data in current fold to check how the model works. Using root mean square error, we can then rank each of the possible parameter values in the parameter space, and the value with the lowest rmse will be the 'optimal' value for that parameter.

\section{Technical Implementation}
** SCREENSHOTS NEEDED **

This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface. Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggrigation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user. This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the inforation to the user in a format which they will understand.

In an implementation of this, the application will begin by having the user enter the username they've registered on \url{boardgamegeek.com}. Instead of logging the users ratings directly through the application, we will instead leverage the existing infrastructure of boardgame geek. After checking that its a valid username, the application will scour boardgamegeek to see if the database's information is complete: that is, do we have this user's data in the player key? Have they rated games that aren't in our database? Have they rerated games that are in our database? Have they rated games that aren't in our database? .If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game). After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we'll proceed to model fitting.

We first apply the global effects strategy as descibed in the modling section. The optimization of the $\alpha$ parameters will have occred sometime before this and the algoritm will use use those stored values. A weeky or monthly automatic updaing procedure is used to keep these values up to date as more data comes into the system. From the global effects, we then aggregate those based on the classifier aggrigation procedure as described above. These are then stored in the database for future use by the system. After that, before we move onto the k-nearest neighbors implementation, we must first find an optimal value for $k$ using the 10-fold cross validation as described above.

Finally, after the k-fold crossvalidation, we're ready to run the other games through the algorithm to get a predicted score. The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user.

Since we saved the aggregated classifier ratings for the individual user, we can suggest ways for the user to subset this data. By providing the graph below to the player (GET THE GRAPH WORKING), they can see for themselves their not-necessarily obvious preferences in game family or mechanic. They can then search by their preferred mechanic or theme, and get ratings for games that fit that description. Or the user can simply sort by highest rated.

@@

\bibliographystyle{asa}
\bibliography{references}
\def\thebibliography#1{\section*{References\@mkboth
 {REFERENCES}{REFERENCES}}\list
%% default is no labels, for those not using \cite or BibTeX
% undent first line of a reference by using negative \itemindent.
 {}{\setlength{\leftmargin}{3em}\setlength{\labelsep}{0pt}\itemindent=-\leftmargin}
 \def\newblock{\hskip .11em plus .33em minus -.07em}
 \sloppy\clubpenalty4000\widowpenalty4000
 \sfcode`\.=1000\relax}

\def\@biblabel#1{\hfill}

% don't box citations, add space between multiple citations, separate with ;
\def\@citex[#1]#2{\if@filesw\immediate\write\@auxout{\string\citation{#2}}\fi
  \def\@citea{}\@cite{\@for\@citeb:=#2\do
    {\@citea\def\@citea{; }\@ifundefined
       {b@\@citeb}{{\bf ?}\@warning
       {Citation `\@citeb' on page \thepage \space undefined}}%
{\csname b@\@citeb\endcsname}}}{#1}}
